---
title: "Agestam 1985"
author: "Carl Vigren"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

As part of the HUGIN project during the late 1970's and 1980's, Agestam presented in his seminal thesis a growth simulator for mixed stands of Pine, Spruce and Birch in Sweden [@agestam1985].

The HUGIN project included the following models for established stands (\>\>8m mean height), as presented by @ekö1985 pp. 9, table 1.

| Species                                                                                 | Type of model      | Data                                  | Reference             |
|-----------------------------------------------------------------------------------------|--------------------|---------------------------------------|-----------------------|
| Scots Pine                                                                              | Stand model        | The Great Yield Investigation (1971). |                       |
| Norway Spruce                                                                           | Stand model        | The Great Yield Investigation (1971). | @eriksson1976         |
| Mixed stands of Pine, Spruce and Birch                                                  | Stand model        | The Great Yield Investigation (1971). | @agestam1985          |
| Pure and mixed stands of pine, spruce, birch, beech, oak and other broad-leaved trees.  | Single-tree model. | NFI                                   | @söderberg1986        |
| Pure and mixed stands of pine, spruce, birch, beech, oak, and other broad-leaved trees. | Stand model        | NFI                                   | @ekö1985              |
| Lodgepole Pine                                                                          | Stand model        | Permanent plots                       | Hägglund et al (1979) |

: Growth simulators for established forests in the HUGIN-system.

These models followed a general sequence set up by @jonsson1980 .

1.  Simulation of young stand and growth up to 8 metres.

2.  A growth period increment length is chosen, c:a 5 years.

For every period.

i)  Age is updated.

ii) Dominant height is calculated from trajectories, e.g.

The development of young stands is estimated by a separate model, e.g. @elfving1975 .

```{r}
# ## Agestam production table 1.
# library(tidyverse)
# library(forester)
# 
# #Start values.
# # Pure Pine stand, T20, total age 38, 2800 stems/ha.  23.8 m2/ha.
# stems <- 2800
# BA <- 23.8
# age <- 38
# SI_Pine <- 20
# increment_period <- 5
# latitude=64
# 
# #age to breast height
# age_to_breast_height <- forester::Hagglund_age_to_height(latitude = latitude,
#                                  SI = SI_Pine,
#                                  dominant_height_m = 1.31,
#                                  species = "Pinus sylvestris",
#                                  regeneration = "culture")[[2]]
# 
# age_at_breast_height <- age-age_to_breast_height
# 
# 
# ## Height equation
# dominant_height_m <- Hagglund_1974_Sweden_height_trajectories_Pine(age_at_breast_height = 100,
#                                                                    top_height_dm = 200,
#                                                                    regeneration ="culture",
#                                                                    age_2 = age_at_breast_height,
#                                                                    output = "Height")/10
# 
# ## start volume
# volume <- Agestam_1985_volume_Pine_SI(basal_area_Pine_m2_ha = BA,
#                                       SI_Pine_m = SI_Pine,
#                                       age_at_breast_height_Pine = age_at_breast_height,
#                                       stems_Pine_ha = stems,
#                                       latitude = latitude)
# 
# 
# #start mean BA diam cm.
# mean_BA_diam_cm <- forester::basal_area_weighted_mean_diameter_cm(basal_area_m2_ha = BA,stem_count = stems)
# 
# ## start thinning 40% BA above bark.
# BA_thinned <- BA*0.4
# stems_thinned <- BA_thinned/(pi*((mean_BA_diam_cm/2)/100)^2)
# 
# BA <- BA-BA_thinned
# stems <- stems-stems_thinned
# 
# #volume_after_thinning
# volume2 <- Agestam_1985_volume_Pine_SI(basal_area_Pine_m2_ha = BA,
#                                       SI_Pine_m = SI_Pine,
#                                       age_at_breast_height_Pine = age_at_breast_height,
#                                       stems_Pine_ha = stems,
#                                       latitude = latitude)
# 
# #thinned_volume
# thinned_volume <- volume-volume2
# 
# volume <- volume2
# 
# output1 <- data.frame("age"=age,
#                       "age_brh"=age_at_breast_height,
#                       "dominant_height_m"=dominant_height_m,
#                       "stems"= stems,
#                       "stems_thinned"=stems_thinned,
#                       "thinned_volume"=thinned_volume,
#                       "BA_diam_cm_living"=mean_BA_diam_cm,
#                       "BA_diam_cm_self_thinned"=0,
#                       "BA"=BA,
#                       "volume"=volume,
#                       "mortality"=0)
# 
# 
# while(age_at_breast_height<=140){
# 
#   age_at_breast_height <- age_at_breast_height+increment_period
#   age<- age+increment_period
# 
#   dominant_height_m <- Hagglund_1974_Sweden_height_trajectories_Pine(age_at_breast_height = 100,
#                                                                      top_height_dm = SI_Pine*10,
#                                                                      regeneration = "culture",
#                                                                      age_2 = age_at_breast_height ,
#                                                                      output = "Height")/10
# 
#   #Thinning
# 
#   #Default 0.
#   stems_thinned <- 0
#   thinned_volume <- 0
# 
#   if((age_at_breast_height-increment_period)<70 & age_at_breast_height>70){
#     BA_thinned <- BA*0.35
#     stems_thinned <- BA_thinned/(pi*((mean_BA_diam_cm/2)/100)^2)
# 
#     BA <- BA-BA_thinned
#     stems <- stems-stems_thinned
# 
#     #volume_after_thinning
#     volume2 <- Agestam_1985_volume_Pine_SI(basal_area_Pine_m2_ha = BA,
#                                            SI_Pine_m = SI_Pine,
#                                            age_at_breast_height_Pine = age_at_breast_height,
#                                            stems_Pine_ha = stems,
#                                            latitude = latitude)
# 
#     #thinned_volume
#     thinned_volume <- volume-volume2
# 
#     volume <- volume2
#   }
# 
# 
# 
#   #Estimate mortality
# 
#   #Mortality during period
#   mortality_m2_ha <- Agestam_1985_mortality_above_bark_Pine(stems_Pine_ha = stems,
#                                                             basal_area_above_bark_all_species = BA,
#                                                             SI = SI_Pine,
#                                                             age_at_breast_height = age_at_breast_height)*increment_period
# 
#   mean_BA_diam_cm <- forester::basal_area_weighted_mean_diameter_cm(basal_area_m2_ha = BA,
#                                                                     stem_count = stems)
#   self_thinned_stem_diam_cm <- forester::Agestam_1985_mortality_diameter_cm(species = "Pinus sylvestris",
#                                                                      age_at_breast_height = age_at_breast_height,
#                                                                      basal_area_weighted_mean_diameter_cm_living = mean_BA_diam_cm)
# 
#   self_thinned_stems <- round(mortality_m2_ha/((pi*(self_thinned_stem_diam_cm/2)^2)/(100*100)),0)
# 
#   stems_thinned <- stems_thinned+self_thinned_stems
# 
# 
#   #bark subtraction
#   bark_period_start_m2_ha <- Agestam_1985_bark_subtraction_Pine(basal_area_over_bark_Pine = BA,
#                                                           SI_Pine = SI_Pine,
#                                                           latitude = latitude
#                                                           )
# 
# 
#   BA_under_bark_period_start_m2_ha <- BA - bark_period_start_m2_ha
# 
#   #BA growth
#   BA_growth_m2_ha <- Agestam_1985_basal_area_annual_increment_under_bark_Pine(basal_area_Pine_m2_ha_after_thinning = BA_under_bark_period_start_m2_ha,
#                                                            stem_count_Pine = stems,
#                                                            age_at_breast_height_period_start = age_at_breast_height,
#                                                            SI_H100_Pine = SI_Pine,
#                                                            basal_area_other_species_m2_ha_after_thinning = 0,
#                                                            diameter_mean_basal_area_stem_Pine_cm = forester::basal_area_weighted_mean_diameter_cm(basal_area_m2_ha = BA_under_bark_period_start_m2_ha,stem_count = stems),
#                                                            diameter_mean_basal_area_stem_all_species_cm = forester::basal_area_weighted_mean_diameter_cm(basal_area_m2_ha = BA_under_bark_period_start_m2_ha,stem_count = stems) ,
#                                                            increment_period_years = increment_period)*increment_period
# 
#   #Total after growth
#   BA_after_growth_m2_ha <- BA_under_bark_period_start_m2_ha+BA_growth_m2_ha
# 
#   #Add bark
#   BA_after_growth_m2_ha <- Agestam_1985_bark_addition_Pine(basal_area_under_bark_Pine = BA_after_growth_m2_ha,
#                                                            SI_Pine = SI_Pine,
#                                                            latitude = latitude)+BA_after_growth_m2_ha
# 
# 
#   #final Basal Area
#   BA <- BA_after_growth_m2_ha-mortality_m2_ha
# 
#   #new stem count
#   stems <- stems-self_thinned_stems
# 
#   #Volume
#   volume <- Agestam_1985_volume_Pine_SI(basal_area_Pine_m2_ha = BA,
#                                             SI_Pine_m = SI_Pine,
#                                             age_at_breast_height=age_at_breast_height,
#                                             stems_Pine_ha = stems,
#                                             latitude = latitude)
# 
# 
#   output2 <- data.frame("age"=age,
#                         "age_brh"=age_at_breast_height,
#                         "dominant_height_m"=dominant_height_m,
#                         "stems"= stems,
#                         "stems_thinned"=stems_thinned,
#                         "thinned_volume"=thinned_volume,
#                         "BA_diam_cm_living"=mean_BA_diam_cm,
#                         "BA_diam_cm_self_thinned"=self_thinned_stem_diam_cm,
#                         "BA"=BA,
#                         "volume"=volume,
#                         "mortality"=mortality_m2_ha)
# 
#   output1 <- dplyr::bind_rows(output1,output2)
# 
# 
# }
# 
# round(output1,1)
# 
# #Correct values from production table 1, p. 86.
# correct <- data.frame("age_brh"=c(29,34,39,44,49,54,61,71,81,91,101,111,121,131,141),"volume"=c(69,87,107,130,153,177,212,175,216,261,306,352,397,442,486.5))
# 
# #Plot results
# ggplot(output1,aes(x=age_brh,y=volume,linetype="own"))+geom_line()+geom_line(data=correct,aes(x=age_brh,y=volume,linetype="correct"))


```
